---
title: "Cross Validation"
output: github_document
---

```{r, include = F}
library(tidyverse)
library(modelr)
library(mgcv)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)
theme_set(theme_minimal() + theme(legend.position = "bottom"))
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_color_viridis_d
scale_fill_discrete = scale_fill_viridis_d

set.seed(1)
```


# How Do You Choose the Best Model using prediction accuracy as a metric?

## Let's start with simulated data

```{r}

nonlin_df = 
  tibble( 
    id = 1:100, 
    x = runif(100, 0, 1),
    y = 1 - 10 * (x - .3) ^ 2 + rnorm(100, 0, .3)
    )

```


Look at the data
```{r}
nonlin_df %>% 
  ggplot(aes(x = x, y = y)) +
  geom_point()
```


Two ways of doing cross validation: by "hand" & in an "automated" fashion using modelr

## Cross validation -- by "hand"

Get training and testing datasets

```{r}
train_df = sample_n(nonlin_df, size = 80)
test_df = anti_join(nonlin_df, train_df, by = "id") #get all observationin nonlin_df that are not in train_df
```


Fit three models. 

```{r}
linear_mod = lm(y ~ x, data = train_df)
smooth_mod = gam(y ~ s(x), data = train_df) #gam stands for generalized additive model
wiggly_mod  = gam(y ~ s(x, k = 30), sp = 10e-6, data = train_df) #more like you're chasing every single point - overfit/wiggly
```

Can I see what I just did ... 

```{r}

train_df %>% 
  add_predictions(linear_mod) %>%  #add_predictions adds the predicted values from that model
  ggplot(aes(x = x, y = y)) + 
  geom_point() + 
  geom_line(aes(y = pred), color = "red")


train_df %>% 
  add_predictions(smooth_mod) %>%  #add_predictions adds the predicted values from that model
  ggplot(aes(x = x, y = y)) + 
  geom_point() + 
  geom_line(aes(y = pred), color = "red")


train_df %>% 
  add_predictions(wiggly_mod) %>%  #add_predictions adds the predicted values from that model
  ggplot(aes(x = x, y = y)) + 
  geom_point() + 
  geom_line(aes(y = pred), color = "red")


```


## as a side we can use `gather_predictions` instead of doing `add_predicitons` for every.single.model.

```{r}
train_df %>% 
  gather_predictions(linear_mod, smooth_mod, wiggly_mod) %>%  #gather_predictions to 'gather' the predicitons of several models at once. peep the data structure
  ggplot(aes(x = x, y = y)) + 
  geom_point() + 
  geom_line(aes(y = pred), color = "red") + 
  facet_grid(. ~ model)
```


Let's asses the `prediciton accuracy` of these models using RMSE to see how well our models fit the `test` dataset


Insert the `RMSE` function. The lower the `RMSE`, the better. 
```{r}
rmse(linear_mod, test_df)
rmse(smooth_mod, test_df) # makes the best predictions according to the RMSE
rmse(wiggly_mod, test_df)
```

# something to think about - is the RMSE we observed liable to what ended up being the training set vs. testing set? could test for this? 

# Cross Validation using the `modelr` package

```{r}
cv_df = 
  crossv_mc(nonlin_df, 100)
```

What is happening in `cv_df` ^^ 



```{r}
cv_df %>% pull(train) %>% .[[1]] %>% as_tibble()

cv_df %>% pull(test) %>% .[[1]] %>% as_tibble()
```


```{r}

cv_df = 
  cv_df %>% 
  mutate(
    train = map(train, as_tibble),
    test = map(test, as_tibble)
  )

```


Let's try to fit models & get RMSEs for them. 
```{r}

cv_df = 
  cv_df %>% 
  mutate(
    linear_mod = map(.x = train, ~lm(y ~ x, data = .x)),
    smooth_mod = map(.x = train, ~gam(y ~ s(x), data = .x)), 
    wiggly_mod = map(.x = train, ~gam(y ~ s(x, k = 30), sp = 10e-6, data = .x))
  ) %>% 
  mutate(
    rmse_linear = map2_dbl(.x = linear_mod, .y = test, ~rmse(model = .x, data = .y)),
    rmse_smooth = map2_dbl(.x = smooth_mod, .y = test, ~rmse(model = .x, data = .y)),
    rmse_wiggly = map2_dbl(.x = wiggly_mod, .y = test, ~rmse(model = .x, data = .y))
  )
```


What do these results say about model choices? 
## VIOLIN PLOTS of RMSE

```{r}

cv_df %>% 
  select(starts_with("rmse")) %>% 
  pivot_longer(
    everything(),
    names_to = "model",
    values_to = "rmse",
    names_prefix = "rmse_"
  ) %>% 
  ggplot(aes(x = model, y = rmse)) + 
  geom_violin() + 
  labs(
    title = "The Distribution of RMSEs across 100 training/testing splits fitting 3 different models", 
    subtitle = "The linear is the worst; Smooth model does the best"
  )
```


We could also compute RMSE averages

```{r}

cv_df %>% 
  select(starts_with("rmse")) %>% 
  pivot_longer(
    everything(),
    names_to = "model",
    values_to = "rmse",
    names_prefix = "rmse_"
  ) %>% 
  group_by(model) %>% 
  summarize(
    avg_mse = mean(rmse))
```



## Let's try this on a real dataset .. 


```{r}
child_growth_df = 
  read_csv("./data/nepalese_children.csv") %>% 
  mutate(
    weight_cp = (weight > 7) * (weight - 7) #change point models to get something that's initally linear up until 7 then is allowed to bend (?) then be straight again
  )
```


weight vs. arm circumference
```{r}

child_growth_df %>% 
  ggplot(aes(x = weight, y = armc)) + 
  geom_point(alpha = .3)
```

Fit the models I care about.

```{r}

linear_mod = lm(armc ~ weight, data = child_growth_df)
pwlin_mod = lm(armc ~ weight + weight_cp, data = child_growth_df) #piicewise linear model that accounts for the weight_cp variable
smooth_mod = gam(armc ~ s(weight), data = child_growth_df)
```


Which of these is the right model to use for this association? 
```{r}
child_growth_df %>% 
  gather_predictions(linear_mod, pwlin_mod, smooth_mod) %>% 
  ggplot(aes(x = weight, y = armc)) + 
  geom_point(alpha = .3) +
  geom_line(aes(y = pred), color = "red") + 
  facet_grid(. ~ model)
```


Try to understand model fit using cross validation . 

```{r}
cv_df = 
  crossv_mc(child_growth_df, 100) %>% 
  mutate(
    train = map(train, as_tibble),
    test = map(test, as_tibble)
  )

```


See if I can fit the model the splits (i.e., to the test datasets) ... 

```{r}

cv_df = 
  cv_df %>% 
  mutate(
    linear_mod = map(.x = train, ~lm(armc ~ weight + weight_cp, data  = .x)),
    pwlin_mod = map(.x = train,  ~lm(armc ~ weight + weight_cp, data  = .x)),
    smooth_mod = map(.x = train, ~gam(armc ~ s(weight), data = .x))
  ) %>% 
  mutate(
    rmse_linear = map2_dbl(.x = linear_mod, .y = test, ~rmse(model = .x, data = .y)),
    rmse_pwlin = map2_dbl(.x = pwlin_mod, .y = test, ~rmse(model = .x, data = .y)),
    rmse_smooth = map2_dbl(.x = smooth_mod, .y = test, ~rmse(model = .x, data = .y))
  )

```


```{r}
cv_df %>% 
  select(starts_with("rmse")) %>% 
  pivot_longer(
    everything(),
    names_to = "model",
    values_to = "rmse",
    names_prefix = "rmse_"
  ) %>% 
  ggplot(aes(x = model, y = rmse)) + 
  geom_violin() + 
  labs(
    title = "The Distribution of RMSEs across 100 training/testing splits fitting the 3 different models: Linear, Piecewise, Smooth", 
    subtitle = "The Smooth model does the best in terms of prediction; but it might be that the 'interpretation' is better when looking at the piecewise model"
  )
```

