---
title: "Cross Validation"
output: github_document
---

```{r, include = F}
library(tidyverse)
library(modelr)
library(mgcv)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)
theme_set(theme_minimal() + theme(legend.position = "bottom"))
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_color_viridis_d
scale_fill_discrete = scale_fill_viridis_d

set.seed(1)
```


# How Do You Choose the Best Model using prediction accuracy as a metric?

## Let's start with simulated data

```{r}

nonlin_df = 
  tibble( 
    id = 1:100, 
    x = runif(100, 0, 1),
    y = 1 - 10 * (x - .3) ^ 2 + rnorm(100, 0, .3)
    )

```


Look at the data
```{r}
nonlin_df %>% 
  ggplot(aes(x = x, y = y)) +
  geom_point()
```


Two ways of doing cross validation: by "hand" & in an "automated" fashion using modelr

## Cross validation -- by "hand"

Get training and testing datasets

```{r}
train_df = sample_n(nonlin_df, size = 80)
test_df = anti_join(nonlin_df, train_df, by = "id") #get all observationin nonlin_df that are not in train_df
```


Fit three models. 

```{r}
linear_mod = lm(y ~ x, data = train_df)
smooth_mod = gam(y ~ s(x), data = train_df) #gam stands for generalized additive model
wiggly_mod  = gam(y ~ s(x, k = 30), sp = 10e-6, data = train_df) #more like you're chasing every single point - overfit/wiggly
```

Can I see what I just did ... 

```{r}

train_df %>% 
  add_predictions(linear_mod) %>%  #add_predictions adds the predicted values from that model
  ggplot(aes(x = x, y = y)) + 
  geom_point() + 
  geom_line(aes(y = pred), color = "red")


train_df %>% 
  add_predictions(smooth_mod) %>%  #add_predictions adds the predicted values from that model
  ggplot(aes(x = x, y = y)) + 
  geom_point() + 
  geom_line(aes(y = pred), color = "red")


train_df %>% 
  add_predictions(wiggly_mod) %>%  #add_predictions adds the predicted values from that model
  ggplot(aes(x = x, y = y)) + 
  geom_point() + 
  geom_line(aes(y = pred), color = "red")


```


## as a side we can use `gather_predictions` instead of doing `add_predicitons` for every.single.model.

```{r}
train_df %>% 
  gather_predictions(linear_mod, smooth_mod, wiggly_mod) %>%  #gather_predictions to 'gather' the predicitons of several models at once. peep the data structure
  ggplot(aes(x = x, y = y)) + 
  geom_point() + 
  geom_line(aes(y = pred), color = "red") + 
  facet_grid(. ~ model)
```


Let's asses the `prediciton accuracy` of these models using RMSE to see how well our models fit the `test` dataset


Insert the `RMSE` function. The lower the `RMSE`, the better. 
```{r}
rmse(linear_mod, test_df)
rmse(smooth_mod, test_df) # makes the best predictions according to the RMSE
rmse(wiggly_mod, test_df)
```

# thought - is the RMSE we observed liable to what ended up being the training set vs. testing set? could test for this? 
