---
title: "Linear models"
output: github_document
---


```{r, include = F}
library(tidyverse)
library(httr)
library(purrr)
library(p8105.datasets)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)
theme_set(theme_minimal() + theme(legend.position = "bottom"))
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_color_viridis_d
scale_fill_discrete = scale_fill_viridis_d

set.seed(1)
```



## Import data
```{r}
data("nyc_airbnb")

nyc_airbnb = 
  nyc_airbnb %>% 
  mutate(stars = review_scores_location / 2) %>% 
  rename(
    borough = neighbourhood_group,
    neighborhood = neighbourhood) %>% 
  filter(borough != "Staten Island") %>%  # not enough data points in staten island so we take it out 
  select(price, stars, borough, neighborhood, room_type)
```


## Fit a model
```{r}
#Trying to plot a line that goes through stars & price, by borough
nyc_airbnb %>% 
  ggplot(aes(x = stars, y = price, color = borough)) + 
  geom_point()
```


Let's fit the model we care about
```{r}
fit = lm(price ~ stars + borough, data = nyc_airbnb)
```


Let's look at the results
```{r, include = F}
fit

#You get a lot of info with summary()
summary(fit)
summary(fit)$coef
coef(fit)
fitted.values(fit)
residuals(fit)
```


Let's look at the result better ... THE BROOM PACKAGE

```{r}
broom::glance(fit)

broom::tidy(fit) %>% 
  select(-std.error, -statistic) %>% 
  mutate(
    term = str_replace(term, "borough", "Borough: ")
  ) %>% 
  knitr::kable(digits = 3)
```


## Be in control of factors

```{r}
nyc_airbnb = 
  nyc_airbnb %>% 
  mutate(
    borough = fct_infreq(borough), 
    room_type = fct_infreq(room_type)
  )
```


Look at the plot again ... peep the re-ordering that took place

```{r}
nyc_airbnb %>% 
  ggplot(aes(x = stars, y = price, color = borough)) + 
  geom_point()
```



Refit the model with the reordered

```{r}
fit = lm(price ~ stars + borough, data = nyc_airbnb)

broom::tidy(fit)
#Manhattan is now my reference value

broom::glance(fit)
#the AIC & signma & r.squared have stayed the same, the only thing that changed is the reference level
```


## Model Diganostics ... the `modelr` package

```{r}
residuals(fit) 

nyc_airbnb %>% 
  modelr::add_residuals(fit) %>% drop_na(resid) %>% 
  ggplot(aes(x = borough, y = resid)) + 
  geom_violin() + 
  ylim(-500, 1500) #add this line to "zoom in" in your scales


nyc_airbnb %>% 
  modelr::add_residuals(fit) %>% filter(resid > 3000) %>% 
  ggplot(aes(x = stars, y = resid)) +
  geom_point() +
  facet_wrap(. ~ borough)
#the residual values here tell me that the price of that airbnb spot is $9.47 higher than what my model would predict; 35.5 dollars higher; $20.5 lower, etc ... 
#^this makes a lot of sense because all that a residual is is a difference between my observed Y & my predicted Y: residuals =  [Y-Y_hat]
```


## Hypothesis tests

#This does ttests by default
```{r}
fit %>% 
  broom::tidy()
```


What about the statistical significance of `borough` as a whole - a partial f 

```{r}

fit_null = lm(price ~ stars, data = nyc_airbnb)
fit_alt = lm(price ~ stars + borough, data = nyc_airbnb)

#partial f test 

anova(fit_null, fit_alt) %>% 
  broom::tidy()
```


## Nest data, fit models

Let's we're interested in whether there's an association between `stars` & `price`, and does that association vary by `borough` ... & moreover, can we see if `stars` & `room_type` has a bigger effect (think about the word `effect` here) in say Manhattan or Queens etc ...  <- to do this, we'll use `interactions` i.e., ask if the effect of `stars` & `room_type` varies by borough ... as opposed to just adjusting for `stars` & `room_type` when regressing price on `price` & `borough`


```{r}
fit = lm(price ~ stars * borough + room_type * borough, data = nyc_airbnb)

broom::tidy(fit)
```
^ This is pretty formal & also pretty complex



The following is more exploratory but maybe easier to understand. 
```{r}

nyc_airbnb %>% 
  nest(data = -borough) %>% 
  mutate(
    models = map(.x = data, ~lm(price ~ stars + room_type, data = .x)), 
    results = map(models, broom::tidy)
  ) %>% 
  select(-data, -models) %>% 
  unnest(results) %>% 
  filter(term != "(Intercept)") %>% 
  select(borough, term, estimate) %>% 
  pivot_wider(
    names_from = borough,
    values_from = estimate
  )
  
```


Let's nest even more . . .


```{r}
nyc_airbnb %>% 
  filter(borough == "Manhattan") %>% 
  nest(data = -neighborhood) %>% 
  mutate(
    models = map(.x = data, ~lm(price ~ stars + room_type, data = .x)), 
    results = map(models, broom::tidy)
  ) %>% 
  select(-data, -models) %>% 
  unnest(results) %>% 
  filter(str_detect(term, "room_type")) %>% 
  ggplot(aes(x = neighborhood, y = estimate)) + 
    geom_point() + 
    facet_wrap(. ~ term) + 
  theme(axis.text.x = element_text(angle = 70, vjust = 0.5, hjust = 0.5))

```



